### 2014-09-11 A1/writeup/writeup.tex
added writeup stuff

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

\bibliographystyle{abbrv}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{The Effect of Uniqueness on Review Helpfulness}
}
% Insert Author names, affiliations and corresponding author email.
\\
Jack Hessel, 
Matthew Milano, 
Maithra Raghu
\end{flushleft}

\section*{Introduction}
Testing this citation \cite{gilbert2010understanding}.

\section*{Data and Methods}

\section*{Reults}

\bibliography{refs}

\end{document}